---
title: "Rollo_python_scripts"
output: github_document
---


# Finding datasets of interest

```{r, echo - FALSE}
# load necesssary libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(rerddap)
library(plotdap)
```

We can use ERDDAP's built in search function to find datasets that match a keyword

### Find drifter datasets from OSMC

```{r}
# use 'rerddap::ed_search()' to search the osmc ERDDAP for "drifter"
osmc_erddap <- "https://osmc.noaa.gov/erddap/"
df_osmc <- ed_search("drifter", which = "tabledap", url = osmc_erddap)
head(df_osmc$info)
```

### Find ADCP data on the IOOS sensors ERDDAP

```{r}
# use 'rerddap::ed_search()' to search the IOOS Sensor ERDDAP for "adcp"
ioos_sensor_erddap <- "https://erddap.sensors.ioos.us/erddap/"
df_sensor <- ed_search("adcp", which = "tabledap", url = ioos_sensor_erddap)
head(df_sensor$info)
```

### Search with allDatasets

For a more sophisticated search, we can use the `allDatasets` dataset from ERDDAP to examine the primary metadata of all datasets.

### Get metadata for IOOS glider datasets

```{r}
# access from the IOOS Glider DAC using 'rerddap::tabledap()"  the "allDatasets" response - https://gliders.ioos.us/erddap/allDatasets.html
ioos_gliders_erddap <- "https://gliders.ioos.us/erddap"
info <- info("allDatasets", url = ioos_gliders_erddap)
df <- tabledap(info)
# list the column names to keep
keep <- c('institution',
           'minLongitude',
           'maxLongitude',
           'minLatitude',
           'maxLatitude',
           'minAltitude',
           'maxAltitude',
           'minTime',
           'maxTime')
# how to do the select in base R
# df_gliders<- subset(df, select = keep)
#
# use 'dplyr::select()' to select desired columns
df_gliders <- select(df, all_of(keep))
# unhappy if class "tabledap" remains,  remove it
class(df_gliders) <- "data.frame"
# use 'lubridate::as_datetime()' to convert "maxTime" and "minTime" to R datetime objects
df_gliders$maxTime <- as_datetime(df_gliders$maxTime)
df_gliders$minTime <- as_datetime(df_gliders$minTime)
head(df_gliders)

```

### Which gliders have uploaded data in the last 6 hours?

```{r}
# use 'lubridate::now()' to get present time as an R datetime object, create new column wiht deploy time
df_gliders$deploy <- now() - df_gliders$maxTime
# use 'dplyr::filter()' to find deploy time less than 6 hours
latest_gliders <- filter(df_gliders, (deploy < 6.))
head(latest_gliders)
```

### Find datasets from deployments that lasted longer than 90 days

```{r}
# use 'dplyr::filter()' to find deploy time greater than 90 days
longest_gliders <- filter(df_gliders, time_length(maxTime - minTime) > 90)
head(longest_gliders)
```

### Which glider missions extended to a depth of greater than 150 m in seas north of 70 N?

```{r}
# use 'dplyr::filter()' to find deploy observations on the latitude/depth range
deep_gliders <- filter(df_gliders, (maxLatitude >= 70) & (minAltitude < -150))
head(deep_gliders)

```

```{r}
# use 'rerddap::tabledap()' to get from 'profile_id', 'time', 'u' from gilder "unit_595-20180804T0000-delayed'
info <- info("unit_595-20180804T0000-delayed", url = ioos_gliders_erddap)
alaska_glider <- tabledap("unit_595-20180804T0000-delayed",
                          fields = c('profile_id', 'time', 'u'),
                          'time>=2018-09-20',
                          url = ioos_gliders_erddap
                   )
# convert character "NaN" to NA
alaska_glider$u[alaska_glider$u == "NaN"] <- NA
# use 'dplyr::filter()' to remove NA
alaska_glider <- filter(alaska_glider, all_of(!is.na(u)))
# convert alaska_glider$u to numeric values
alaska_glider$u <- as.numeric(alaska_glider$u )
# get parameter long_name and units from the 'info()' call
ylab = paste(info$alldata$u$value[9], info$alldata$u$value[13])
# plot using "ggplot()"
ggplot()  + 
    geom_point(data = alaska_glider , aes(y = u, x= 1:(length(u))), size = .5) +
    xlab('Profile') + ylab(ylab) 

```

------------------------------------------------------------------------

### References

IOOS ERDDAP https://gliders.ioos.us/erddap

More info on using ERDDAP's inbuilt search https://ioos.github.io/erddapy/01b-tabledap-output.html
